---
title: "Statistical analyses and visualization in R:I Assingment 1"
author: "Yuka Tatsumi"
date: "2021-10-15"
output: 
  word_document: 
    fig_height: 7
    fig_width: 7
---

```{r, message=FALSE, warning=FALSE, include=FALSE}
library(Rcmdr) #this chunk start RCommander 
library(ggplot2) #this chunks starts ggplot2
library(flextable)
library(faraway)
library(car)
library(dplyr)
#library(abind, pos=17)
#library(e1071, pos=18)
#if you have used other packeges you may need to include them here
```


```{r, message=FALSE, warning=FALSE, include=FALSE}
#include a chunk that opens your indata
Dataset1 <- 
  read.table("C:/Users/y_tat/Downloads/statsitics_uni/söder/Assignment1/svar/data.txt",
   header=TRUE, stringsAsFactors=TRUE, sep="\t", na.strings="NA", dec=",", 
  strip.white=TRUE)

Dataset2 <- 
  read.table("C:/Users/y_tat/Downloads/statsitics_uni/söder/Assignment1/svar/data2.txt",
   header=TRUE, stringsAsFactors=TRUE, sep="", na.strings="NA", dec=",", 
  strip.white=TRUE)

Dataset3 <-
readXL("C:/Users/y_tat/Downloads/statsitics_uni/söder/Assignment1/Lander_HousingNew.xlsx",
  
rownames=FALSE, header=TRUE, na="", sheet="Sheet1", stringsAsFactors=TRUE)

Dataset4 <- 
  read.table("C:/Users/y_tat/Downloads/statsitics_uni/söder/Assignment1/Housepricing_sweden.txt",
   header=TRUE, stringsAsFactors=TRUE, sep="\t", na.strings="NA", dec=",", 
  strip.white=TRUE)

```


## Result

### Part 1. Correlation between GDP and access to personal computers in the year of 2005

### Question 1. Is there a relationship between GDP/capita and the number of personal computers per 100 individuals measured in 2005?

#### Median, mean and SD

The median, mean and SD for GDP per Capita are as follows according to the calculation with R.

GDP Capita  
Median  :2196.2, Mean:7963.5, SD:12328.80814


Regarding the number of personal computers per 100(PC_PER.100), the median, mean and SD are as follows.

Number of personal computers per 100(PC_PER.100)  
Median:5.900, Mean:15.781, SD:22.11974


**Summaries: Active data set**
```{r}
summary(Dataset1) 

```



**Summaries: Numerical summaries**
```{r}
numSummary(Dataset1[,
  c("GDPperCAPITA", 
  "PC_PER.100"), drop=FALSE], 
  statistics=c("mean", "sd", 
  "IQR", "quantiles"), 
  quantiles=c(0,.25,.5,.75,1))

```


####  A plot to check normal distribution for variables and Choose the appropriate test.


The histograms for both GDP per capita and for the number of personal computers per 100 individuals below show that both variables are not normally distributed.

**Histogram to check normal distribution for GDP per Capita**

```{r}
ggplot(Dataset1) +
  aes(x = GDPperCAPITA) +
  geom_histogram(bins = 30L, fill = "#112446") + 
  labs(x = "GDP per Capita", y="frequency", title ="Histogram:GDP per capita", caption = "Data source: Gapminder(https://www.gapminder.org/data/)" ) +
  theme_classic()+ 
  theme(plot.title = element_text(hjust = 0.5))  
  

```


** Histogram to check normal distribution for PC per 100 people **

```{r}

ggplot(Dataset1) +
  aes(x = PC_PER.100) +
  geom_histogram(bins = 30L, fill = "#112446") +
  labs(x = "PC per 100 people", y="frequency", title="Histogram:PC per 100 people", caption="Data source: Gapminder(https://www.gapminder.org/data/)" ) +
  theme_classic()+
  theme(plot.title = element_text(hjust = 0.5))  

```


#### Choice of appropriate test and motivation 

According to the histograms above, both of the variables are not normally distributed. Therefore, Spearman Rank-Order correlation is a test to be used because Pearson product-moment correlation requires normal distribution of the variables. 


####  Result and analysis of the statistic values and a scatter plot

**Result and analysis of the statistic**

According the Spearman Rank-Order correlation's test result below, correlation coefficient is 0.8547637, degree of freedom= 153  and P value is 2.2e-16. The p value suggests that correlation coefficient 0.8547637 is statistically significant. It implies that there is a correlation between GDP per capita and number of personal computers per 100 individuals in 2005. (95% confidence)

```{r}
with(Dataset1, 
  cor.test(GDPperCAPITA, PC_PER.100, 
  alternative="two.sided", 
  method="spearman"))
```




**A scatter plot **

The scatter plot below shows the positive linear relationship between between GDP per capita and number of personal computers per 100 individuals in 2005. It indicates that the higher the GDP per capita in the country is, the higher the percentage of PC per 100 individuals is.

However, there are several outliers especially around higher GDP capita such as Luxembourg and Monaco.

Luxembourg	GDP: 51927,36096, PC per 100:62,38  
Monaco	    GDP: 81827,95612, PC per 100: 42,55  



```{r}

ggplot(Dataset1) +
 aes(x = GDPperCAPITA, y = PC_PER.100) +
 geom_point(shape = "circle", size = 1.5, colour = "#112446") +
 labs(x = "PC per 100 people", y = "GDP per capita",title="PC per 100 people vs GDP per capita",caption="Data source: Gapminder(https://www.gapminder.org/data/)") +
theme_classic()+
theme(plot.title = element_text(hjust = 0.5))

```





### Part 2 Regression analysis 
### Question 2 Has the electricity generation per capita in China increased from 1990 to 2005?


#### Regression Analysis

#### Making a linear regression 
The following linear regression model is created, using R. 

```{r}
RegModel.1 <- lm(EL_China~Year, data=Dataset2)
RegModel.1 %>% as_flextable()

```


#### Checking if the model is well adapted to the data set.

Graph for "Residuals vs Fitted" in the basic diagnostic plots below shows U shape patterns and does not show horizontal line. This means weak linear relationships in this model. "Normal Q-Q shows plots" follows roughly reference line, which means the residual are roughly normally distributed. "Scale-location plots" do not show a horizontal line and plots spread unequally, which indicates heteroscedasticity.  "Resiudals vs Leverage" chart shows that there are no outliers that may affect the model as all plots are within Cook's distance line.



**Basic Diagnostic plots **
```{r include=FALSE}

oldpar <- par(oma=c(0,0,3,0), 
  mfrow=c(2,2))
```


```{r}
par(mfrow=c(2,2))
plot(RegModel.1)
par(mfrow=c(1,1))

```


  
**Scatterplot**

The scatterplot (year and electricity generation per capita in china) below does not show complete linear. There is a slight curve around year 2000, which shows higher growth rate of electricity generation after 2000.


```{r include=FALSE}
par(oldpar)
```

```{r}

ggplot(Dataset2) +
  aes(x = Year, y = EL_China) +
  geom_point(shape = "circle", size = 1.5, colour = "#112446") +
  labs(x = "Year", y = "Electricity generation per capita", title = " Electricity generation per capita (kilowatt-hours) in China", caption="Data source: Gapminder(https://www.gapminder.org/data/)") +
  theme_classic()+
  stat_smooth(method="lm", formula=y~x, geom="smooth")+
  theme(plot.title = element_text(hjust = 0.5))  


```


####  Transform the response variable, using log 10. 
The evaluation of basic diagnostic plots and the scatterplot (year and electricity generation per capita in China) above shows that the regression model(RegModel.1) that is created above needs to be improved. In order to improve the regression model, log 10 transformation is used to create the new model as below.


```{r}
RegModel.2 <- 
  lm(log10(EL_China)~Year, 
  data=Dataset2)
```



According to the test result of the model after log 10 transformation(RegModel.2)below, Adjusted R-squared has improved from 0.8706 to 0.9689. It indicates that this model is better fit than the first model.

```{r}
RegModel.2 %>% as_flextable()

```



**Diagnostic Diagram for RegModel.2**  

Graph for "Residuals vs Fitted" in the basic diagnostic plots still shows U shape patterns but less than the previous model (RegModel.1). "Normal Q-Q shows plots" follows roughly reference line, which means the residual are roughly normally distributed. In "Scale-location" chart, plots do not show a complete horizontal lines that indicates homoscedasticity but closer to horizontal than the previous model. "Residuals vs Leverage" chart shows that there are no outliers that may affect the model as all plots are within Cook's distance line.


```{r include=FALSE}
oldpar <- par(oma=c(0,0,3,0), 
  mfrow=c(2,2))
```


```{r}
par(mfrow=c(2,2))
plot(RegModel.2)
par(mfrow=c(1,1))
```



#### Transforming the response variable, using square root. 
In order to search for a better fit regression model, the response model is transformed, using square root. 

According to the test result below, Adjuststed R-squared (0.9278) in the model (RegModel.3) has not improved from the previous model(RegModel.2, Adjusted R-squared (0.9689)).  It suggests that the previous model(RegModel.2) is better fit than the model(Regmodel.3).


```{r}
RegModel.3 <- 
  lm(sqrt(EL_China)~Year, 
  data=Dataset2)
```

```{r}
RegModel.3 %>% as_flextable()

```


Graph for "Residuals vs Fitted" in the basic diagnostic plots below shows U shape patterns. "Normal Q-Q", plots follows reference line. In "Scale-location" chart,plots do not show complete horizontal lines. "Residuals vs Leverage" chart shows that there are no outliers that may affect the model as all plots are within Cook's distance line.

```{r}
par(mfrow=c(2,2))
plot(RegModel.3)
par(mfrow=c(1,1))
```


```{r include=FALSE}
par(oldpar)
```

```{r}

ggplot(Dataset2) +
  aes(x = Year, y = sqrt(EL_China)) +
  geom_point(shape = "circle", size = 1.5, colour = "#112446") +
  labs(x = "Year", y = "Sqrt(Electricity generation per capita)",title = "Sqrt(Electricity generation per capita in China)", caption = "Data source: Gapminder(https://www.gapminder.org/data/)" )+
  theme_classic()+
  stat_smooth(method="lm", formula=y~x, geom="smooth")+
  theme(plot.title = element_text(hjust = 0.5))  

```




#### Comparison of all the different models

Comparison of Adjusted R2 value of the 3 models suggested that RegModel.2 are best model to use. (R-squared:  0.9707,	Adjusted R-squared:  0.9689 F-statistic: 562.4 on 1 and 17 DF,  p-value<0.01)

Therefore, RegModel.2 (b(regression coefficient): 0.036687, SEb = 0.001547, t=23.71, p<0.01 df=17) below is the best fit among 3 models. 



```{r}
RegModel.2 %>% as_flextable()

```



```{r}

ggplot(Dataset2) +
  aes(x = Year, y = log10(EL_China)) +
  geom_point(shape = "circle", size = 1.5, colour = "#112446") +
  labs(x = "Year", y = "Log10(Electricity generation per capita)", title = "Log10(Electricity generation per capita in China)", caption = "Data source: Gapminder(https://www.gapminder.org/data/)")+
  theme_classic()+
  stat_smooth(method="lm", formula=y~x, geom="smooth")+
  theme(plot.title = element_text(hjust = 0.5))  



```


### Part 3. Testing differences between groups

### Question 3. Is there a difference in income between the New York districts, Manhattan and Brooklyn?


####  Histogram and appropriate test to use
According to the histogram below, variable income is not normally distributed. Therefore, Wilcoxon's rank-sum test is more appropriate to use.


```{r}
options(scipen=999)
ggplot(Dataset3) +
  aes(x = Income) +
  geom_histogram(bins = 30L, fill = "#112446") +
  labs(x = "Income", y="frequency", title="Income in Manhattan and Brooklyn" ,caption="Lander (2019) (https://www.jaredlander.com/datasets/)") +
  theme_classic()+
  theme(plot.title = element_text(hjust = 0.5))  
```


#### 2. Run Wilcoxon's rank-sum test
The result of Wilcoxon's rank-sum test is as below (W = 223, p-value = 0.00001019, df= 21). The p-value( 0.00001019) indicates that there are statistically significant differences in income between Manhattan and Brooklyn. 



```{r}
wilcox.test(Income ~ Boro, alternative="two.sided", data=Dataset3)
```


**Calculation of median, mean, SD**

**Median **

There is a big difference in median of income between in Manhattan and Brooklyn in the dataset. The median in Manhattan is much higher than the one in Brooklyn. 

```{r}
Tapply(Income ~ Boro, median, na.action=na.omit, data=Dataset3)
```


```{r}

with(Dataset3, tapply(Income, Boro,numSummary))

```

**MEAN**

There is a big difference in mean of income between Manhattan and Brooklyn in the dataset.The mean in Manhattan is much higher than the one in Brooklyn.

```{r}
with(Dataset3,tapply(Income,Boro, mean))

```

**SD**

There is a big difference in standard deviation of income between Manhattan and Brooklyn in the dataset. The standard deviation in Manhattan is much bigger than the one in Brooklyn.

```{r}
with(Dataset3,tapply(Income,Boro, sd))

```


#### Graphs to visualize findings

The boxplot below shows that the median of income in Manhattan is higher than in Brooklyn and there is a big difference. However, the difference in upper quartiles as well as the difference in maximum between Brooklyn and Manhattan is much bigger than the difference of median. Additionaly, the dataset for Manhattan is right skewed with a large variance/spread and several extreme outliers exist around higher income in Manhattan. 


```{r}
options(scipen=999)
ggplot(Dataset3) +
  aes(x = Boro, y = Income) +
  geom_boxplot(shape = "circle", fill = "#112446") +
  labs(x = "Boro", y = "Income", title="Income in Manhattan and Brooklyn" ,caption="Lander (2019) (https://www.jaredlander.com/datasets/)") +
  theme_classic()+
  theme(plot.title = element_text(hjust = 0.5))  

```





#### Interpretation of the results
According to the the result of Wilcoxon's rank-sum test mentioned above(W = 223, p-value = 0.00001019 and df= 21),
there are significant difference in income between Manhattan and Brooklyn.


### Question 4 Are there differences in house pricing (SEK/m2) in Sweden between 2016 and 2017?


####  Histograms and appropriate test to use

According to the histograms below, both variable(housing price in Sweden in 2016 and 2017) are not normally distributed. Therefore, wilcoxon singed rank exact test is appropriate to use. 



```{r}
ggplot(Dataset4) +
  aes(x = X2016_sek_sqrm) +
  geom_histogram(bins = 30L, fill = "#112446") +
  labs(x = "Housing Pirce", y = "Frequency", title="Housing Pirce in Sweden in 2016", caption="Svensk mäklarstatistik (https://www.maklarstatistik.se/)") +
  theme_classic()+
 theme(plot.title = element_text(hjust = 0.5))  

```


```{r}

ggplot(Dataset4) +
  aes(x = X2017_sek_sqrm) +
  geom_histogram(bins = 30L, fill = "#112446") +
  labs(x = "Housing Pirce in Sweden", y = "frequency", title="Housing Pirce in Sweden in 2017", caption="Svensk mäklarstatistik (https://www.maklarstatistik.se/)") +
  theme_classic()+
  theme(plot.title = element_text(hjust = 0.5)) 
  

```





#### Perform Wilcoxon signed rank exact test

The paired samples test is appropriate  because both variables(X2016_sek_sqrm, x2017_sek_sqrm) come from the same sample(same month in different year) and they are not totally independent from each other. 


P-value from the test is 0.006836 with V=6 and degree of freedom 11. This suggests that the difference between 2016 and 2017 is statistically significant. 


```{r}
with(Dataset4, median(X2016_sek_sqrm - X2017_sek_sqrm, na.rm=TRUE)) 
  # median difference
with(Dataset4, wilcox.test(X2016_sek_sqrm, X2017_sek_sqrm, 
  alternative='two.sided', paired=TRUE))
```


#### A Plot to dispalay the diffrence 
The plot below also shows that there is a difference between housing price in Sweden in 2016 and 2017 and that housing price in 2017 is higher than 2016. But it also shows that 2 outliers (november and december) in 2017 exists.


```{r}

sqrm<-append(Dataset4$X2016_sek_sqrm,Dataset4$X2017_sek_sqrm)
label<-factor(c(rep("2016",12),rep("2017",12)))

```


```{r}
Dataset4a<-data.frame(label, sqrm)
ggplot(Dataset4a) +
  aes(x = label, y = sqrm) +
  geom_boxplot(shape = "circle", fill = "#112446") +
  labs(x = "year", y = "House Price in Sweden", title="Housing Pirce in Sweden", caption="Svensk mäklarstatistik (https://www.maklarstatistik.se/)") +
  theme_classic()+
  theme(plot.title = element_text(hjust = 0.5)) 



```



## Discussion

#### Question 1 Is there a relationship between GDP/capita and the number of personal computers per 100 individuals measured in 2005?

The Spearman Rank-Order correlation's test result shows that, there is correlation between GDP per capita and number of personal computers per 100 individuals in 2005. (95% confidence)  It indicates that the country who has higher GDP per capita tend to have a higher percentage of PC per 100 individuals. However, there are some exceptions  especially around higher GDP per capita such as Luxembourg and Monaco, which suggests that there might be another factor that can influence the percentage of PC per 100. This can be investigated further. 

#### Question 2 Has the electricity generation per capita in China increased from 1990 to 2005?

According to comparison between the 3 linear regression models based on	Adjusted R-squared, it can be concluded that Regmodel2 with log 10 transformation of dependent variable(electricity generation) are best fit. This implies that electricity generation per capita in China increased exponentially from 1990 to 2005. It increased more rapidly after around 2000. 


#### Question 3 Is there a difference in income between the New York districts, Manhattan and Brooklyn?

The results of Wilcoxon's rank-sum test indicates that there are statistically significant difference in income between Manhattan and Brooklyn and Income in Manhattan is more likley to be higher than Brooklyn. However, several extreme outliers (high income) and a large variance with right skewed are found in data of Manhattan, which means that several very high incomes in Manhattan increased the average income in Manhattan. 


 
#### Question 4. Are there differences in house pricing (SEK/m2) in Sweden between 2016 and 2017?

The result suggests that the difference in house pricing between 2016 and 2017 is statistically significant. It indicates that there are differences in house pricing (SEK/m2) in Sweden between 2016 and 2017. It can be said that the housing price in Sweden has increased between 2016 and 2017 in total(Jan-Oct). However, it is good to notice that the dataset shows that the housing price decreased in Oct, Nov and December in 2017 and the housing price in November and December in 2017 was lower than the same months in 2016. It means that there was a relatively large decline in housing price from October to December 2017 in contrast to the constant increase that was seen before.